Reduced growth due to belowground sink limitation is not explained by reduced photosynthesis

### 2016-04-13 manuscript/manuscript_pve.Rmd
manuscript edits BM

\
Word counts
\
Main text:
\
Introduction:
\
Materials and methods:
\
Results:
\
Discussion:
\
Acknowledgements:
\
7 figures (6 color), 3 tables
\
Supporting information: 1 table

@@

- Container volume negatively affected aboveground growth throughout the experiment, yet, mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containered seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. Although belowground sink limitation reduced *A*~n~, the model concluded that these reductions were not large enough to explain observed growth responses.

- As *A*~n~ and growth were not tightly coordinated, an unaccounted-for pool of photosynthetic C resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

\
Manipulations of container size were expected to induce a belowground sink limitation compared to free seedlings, resulting in growth reductions in the smallest containers. Specifically, reductions in biomass partitioning to root mass, relative to tree size, were predicted to occur with decreasing rooting volume and soil nutrients in container treatments. We also expected a correlation between carbohydrate accumulation and photosynthetic capacity as a function of soil volume, as reduced sink strength would lead to accumulation of leaf non-structural carbohydrates and down regulation of *A~n~*. Taking into account measured rates of photosynthesis, we expected biomass predictions from the growth model to correspond to observed seedling mass

@@

\
Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil (described above). Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

##Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. After four months, a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

@@

##Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of nitrogen (N) content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of N. Nitrogen concentrations ([N]) of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentage of N in the sample was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in @mitchell2013drought. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, was calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf N content. Similarly, TNC-free leaf N (N~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC content.

##Seedling growth model
We developed a simple seedling growth model that utilized leaf *A*~n~ rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~0, stem~0~ and root~0~) and a starting leaf area (L~0~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (*G*), from @makela1997carbon, is then given by

@@

where *R* is published or measured tissue respiration rates (g C g mass^-1^ d^-1^) of stems (s), fine roots (fr) and coarse roots (cr) and *M* is the modelled standing biomass for each tissue component (g). Leaf respiration is included in the calculation of *P*. The change in each tissue component biomass (*M*) on daily time steps, including leaves, stems fine roots and coarse roots, is given by

@@

\
We then used this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. First, a default model with mean photosynthetic parameters from the free seedlings was optimized to estimate a leaf mass fraction (LMF) which correctly predicted both the leaf mass and total biomass of the harvested free seedling (*M*~free~). This optimized LMF was then used to constrain model runs with treatment-specific *P*, while keeping woody tissue respiration and tissue turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Cumulative net leaf C gain for each treatment was equal to the sum of each value of *P* over 120 days and final seedling C was assumed to equal half of the final dry mass for both modeled and observed seedlings. Seedling biomass production was compared between model output and harvested seedlings with treatment-specific mean *P* by first scaling values to the free seedling control.

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N content and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

##Modelling seedling biomass
Optimization of the default model (*M*~free~) against data yielded an estimate of LMF of 25.7 %; with this LMF, the model predicted mean harvest total biomass of free seedlings within 0.1 %. Using this optimized LMF, the modeled total biomass of seedlings for each soil volume treatment (*M*~pots~) was on average 23&pm;2.4 g C more than measured seedling biomass when compared against predicted total net leaf C gain (Figure `r massmodelfig`a). As a result, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the soil volume treatments when non-photosynthetic parameters were kept constant (Figure `r massmodelfig`b). As a result, the observed reductions in leaf *A* with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the  container size treatments.

@@

##Reductions in growth and physiology under sink limitation
Soon after seedlings became established, both height and diameter growth were negatively affected by decreasing soil volume. Low growth led to large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter, Bühler, et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, we observed a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

\
With respect to nutrient availability, reductions in *A*~n~ were positively correlated with decreases in leaf [N] and leaf N was considerably reduced for seedlings in containers. Since reductions in leaf N were detected with TNC-free leaf mass, it is likely that nutrient uptake was affected in the smallest containers instead of TNC dilution in leaves. It it possible that physical root restriction or decreased supply reduced seedling N uptake in small containers. Although root N was on average higher in free seedlings at the end of the experiment, it was not consistently higher than the smallest container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N in free seedlings, but this effect is unknown. Soil N also declined evenly across all treatments, providing no clear evidence for decreased supply between free seedlings and seedlings in containers. In these already low fertility soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the effects of belowground sink limitation on *A*~n~ of seedlings in containers was evident throughout the experiment.

@@

##Do reductions in photosynthesis explain reductions in seedling growth?
Our model used a simple approach to drive seedling growth with measured reductions in leaf *A*~n~ , via soil volume effects, while treating C use efficiency, respiration and C allocation as fixed processes. Contrary to expectation, the model consistently overestimated seedling growth in containers when parameterized with an optimized LMF for free seedlings. Although reductions in both *A*~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter et al. [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. These findings are important as this model reflects classical approaches in tree growth and production modelling that are driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible that the overestimation of growth was due to an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *Vc*~max~, *R* and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in carbon-balance driven  plant growth models [@lacointe2000carbon]. To address this issue, we utilized the flexibility of this model to test plausible fates of the pool of simulated non-biomass C unaccounted for with observed mass balance. Similar to Lohier et al. [-@lohier2014explaining] we manipulated processes contributing to modeled seedling C mass balance, including changes to leaf C allocation or non-leaf tissue respiration, to quantitatively test their respective influences on model predictions.

\
Using measured LMF from the harvest, instead of the optimized seedling control (*M*~S1~), improved biomass predictions and provided insight into how sink limitation can impact leaf C allocation beyond *A*~n~. The sensitivity of the model to shifts in LMF could represent leaf loss throughout the experiment that could not be explicitly quantified in this field study. As TNC accumulation can lead to accelerated leaf senescence [@paul2001sink], this could explain the observed decline in total leaf area of seedlings in small containers. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated to growth, respiration and C storage in leaves. It will be the interactions between these two components that will determine the total C gain available for plant growth.

\
**this section needs to address why different, implications for models (what they do and need to do)*

Although the fraction of photosynthate used in respiration is known to vary among species and is sensitive to changes in growth rates [@Lambers2008], it is often conisdered a static parameter in process based growth models.

These results imply that using fixed rates of respiration in models likely underestimates plant responses to sink limitation. Thus, we agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models.

A lack of knowledge regarding how respiration rates of individual tissues, within a single plant, maybe be differentially affected by environmental change.

This is noteworthy, as C balance is a delicate equilibrium between fluxes of *A*~n~ and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions [@valentini2000respiration].

Our findings reveal that a combination of different mechanisms, beyond *A*~n~, are likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate growth will undoubtedly shift across different experimental manipulations and plant species.

Empirical data in this study point to *A*~n~ driving reductions in growth, however, modelling result highlight the importance of investigating other components of C balance.

*growth limitation affects both photosynthesis and other components of the C balance*

##Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, when compared with naturally planted seedlings. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants is seldom studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools such as TNC or respiration were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant, changes in *A*~n~ were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control *A*~n~ through the demand for C. Although this may be true, our results imply that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

# Tables {.unnumbered}
**Table 1**.Responses of plant and leaf characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment. Seedling mass and leaf $\delta$^13^C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The volume effect *P* value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

dust(table1)%>%
  sprinkle_colnames("Volume (L)", 
                        "Seedling mass (g)", 
                        "SLA~TNC-free~ (m^2^ kg^-1^)",
                        "Leaf Starch (%)", 
                        "Leaf Sugars (%)", 
                        "Leaf Nitrogen~TNC-free~ (%)",
                        paste0("Leaf ", "$\\delta$","^13^C"," (\u2030)")) %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

# Acknowledgements
We thank Burhan Amiji for his outstanding technical assistance. We would like to thank the many Hawkesbury Institute for the Environment staff and students who helped during the experimental harvest. This research was funded by a PhD Scholarship from the Hawkesbury Institute for the Environment (Western Sydney University).

>>>>

\
Word counts
\
Main text:
\
Introduction:
\
Materials and methods:
\
Results:
\
Discussion:
\
Acknowledgements:
\
6 figures (6 color), 3 tables
\
Supporting information: 1 table

@@

- Container volume negatively affected aboveground growth throughout the experiment, yet, mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containered seedlings, and was related to reduced leaf nitrogen content and starch accumulation. Although belowground sink limitation reduced *A*~n~, the model concluded that these reductions were too small to explain observed growth responses.

- As *A*~n~ and growth were not tightly coordinated, an unaccounted-for pool of photosynthetic C resulted in seedlings with soil volume restriction. These results highlight the need to utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

\
Manipulations of container size were expected to induce a belowground sink limitation compared to free seedlings, resulting in growth reductions in the smallest containers. Specifically, reductions in biomass partitioning to root mass, relative to tree size, were predicted to occur with decreasing rooting volume and soil nutrients in container treatments. We also expected a correlation between carbohydrate accumulation and photosynthetic capacity as a function of soil volume, as reduced sink strength would lead to accumulation of leaf non-structural carbohydrates and down regulation of *A~n~*. Taking into account measured rates of photosynthesis, we expected biomass predictions from the growth model to correspond to observed seedling mass.

@@

\
Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil. Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

##Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. After four months, a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as matted roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

@@

##Leaf, root and soil chemistry
Leaves used in each gas exchange measurements and subsamples of harvested roots were dried to a constant mass and milled for analysis of nitrogen (N) content, $\delta$^13^C, and total non-structural carbohydrates (TNC). Pre-planting soil samples (n=6) and subsamples of soil from each container following harvest were sieved to remove organic material, air dried and milled for analysis of N. Nitrogen concentrations of leaf and soil samples were determined using a Carlo Erba CE1110 elemental analyzer with thermal conductivity and mass spectromic detection (of N~2~ and CO~2~). The percentange of N in samples was calculated by comparison with certified standards. Leaf $\delta$^13^C was analyzed  with an Delta V Advantage coupled to a Flash HT and Conflo IV isotope ratio mass spectrometer. Leaf samples were flash combusted at 1000&deg;C to convert to CO~2~, feed to the mass spectrometer and isotopic signatures are reported relative to the standard Vienna Pee Dee Belemnite scale.

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in @mitchell2013drought. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, was calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass. Similarly, TNC-free leaf N content (N~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC content.

##Seedling growth model
We developed a simple seedling growth model that utilized leaf *A*~n~ rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~0~, stem~0~ and root~0~) and a starting leaf area (L~0~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (*G*), from @makela1997carbon, is then given by

@@

where *R* is published or measured tissue respiration rates (g C g mass^-1^ d^-1^) of stems (s), fine roots (fr) and coarse roots (cr) and *M* is the modelled standing biomass for each tissue component (g). Leaf respiration is included in the calculation of *P*. The change in each tissue component biomass (*M*) on daily time steps, including leaves, stems, fine roots and coarse roots, is given by

@@

\
We then used this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. First, a default model with mean photosynthetic parameters from the free seedlings was optimized to estimate a leaf mass fraction (LMF) which correctly predicted both the leaf mass and total biomass of the harvested free seedlings (*M*~free~). This optimized LMF was then used to constrain model runs with treatment-specific *P*, while keeping woody tissue respiration and tissue turnover parameters constant, to determine if changes in leaf *A*~n~ alone could predict biomass of seedlings in containers (*M*~pots~). Cumulative net leaf C gain for each treatment was equal to the sum of each value of *P* over 120 days and final seedling C was assumed to equal half of the final dry mass for both modeled and observed seedlings. Seedling biomass production was compared between model output and harvested seedlings with treatment-specific mean *P* by first scaling values to the free seedling control.

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

##Modelling seedling biomass
Optimization of the default model (*M*~free~) against data yielded an estimate of LMF of 25.7 %; with this LMF, the model predicted mean harvest total biomass of free seedlings within 1.0 %. Using this optimized LMF, the modeled total biomass of seedlings for each soil volume treatment (*M*~pots~) was on average 11&pm;2.3 g C more than measured seedling biomass when compared against predicted total net leaf C gain (Figure `r massmodelfig`a). As a result, seedling C mass was overestimated by an average of 45&pm;6.2 % in modeled seedlings across the soil volume treatments when non-photosynthetic parameters were kept constant (Figure `r massmodelfig`b). As a result, the observed reductions in leaf *A* with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the container size treatments.

@@

##Reductions in growth and physiology under sink limitation
Soon after seedlings became established, both height and diameter growth were negatively affected by decreasing soil volume. Low growth led to large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, we observed a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

\
With respect to nutrient availability, reductions in *A*~n~ were positively correlated with decreases in leaf N~f~ and leaf N~f~ was considerably reduced for seedlings in containers. Since reductions in leaf N~f~ were detected with TNC-free leaf mass, it is likely that nutrient uptake was affected in the smallest containers instead of TNC dilution in leaves. It it possible that either physical root restriction or decreased supply reduced seedling N uptake in small containers. Although root N was on average higher in free seedlings at the end of the experiment, it was not consistently higher than the smallest container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N~f~ in free seedlings, but this effect is unknown. Soil N also declined evenly across all treatments, providing no clear evidence for decreased supply between free seedlings and seedlings in containers. In these already low fertility soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N~f~, the effects of belowground sink limitation on *A*~n~ of seedlings in containers was evident throughout the experiment.

@@

##Do reductions in photosynthesis explain reductions in seedling growth?
Our model used a simple approach to drive seedling growth with measured reductions in leaf *A*~n~ , via soil volume effects, while treating C use efficiency, respiration and C allocation as fixed processes. Contrary to expectation, the model consistently overestimated seedling growth in containers when parameterized with an optimized LMF for free seedlings. Although reductions in both *A*~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter et al. [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. These findings are important as this model reflects classical approaches in tree growth and production modelling that are driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible that the overestimation of growth was due to an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *Vc*~max~, *R* and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in carbon-balance driven plant growth models [@lacointe2000carbon].

\
Focusing only on empirical results in this study points to *A*~n~ driving reductions in seedling growth. Our modelling results reveal that other components of the C balance, beyond *A*~n~, are required to explain the observed seedling biomass response to sink manipulation. This discrepancy is noteworty as partial accounting of the different mechaninsms involved in plant carbon balance can lead to erroneous conclusions [@valentini2000respiration]. For example, the fraction of photosynthate used in respiration is known to vary depending on species and local environmental conditions [@Lambers2008], yet is often considered a static parameter in process based growth models. The use of fixed rates of respiration in our model likely underestimates seedling responses to sink limitation. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of *A*~n~ and the fate of C allocated to growth, respiration and C storage in leaves. However, the degree to which these mechanisms regulate the C available for growth will undoubtedly shift across different species and environmental conditions.

##Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, when compared with naturally planted seedlings. This finding is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants is seldom studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant, changes in *A*~n~ were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control *A*~n~ through the demand for C. Although this may be true, our results imply that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate.

Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents.

The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

# Tables {.unnumbered}
**Table 1**.Responses of plant and leaf characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment. Seedling mass and leaf $\delta$^13^C values are from final harvest. Values of SLA~f~, leaf starch, leaf sugars and leaf N~f~ represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The volume effect *P* value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

dust(table1)%>%
  sprinkle_colnames("Volume (L)", 
                        "Seedling mass (g)", 
                        "SLA~f~ (m^2^ kg^-1^)",
                        "Leaf Starch (%)", 
                        "Leaf Sugars (%)", 
                        "Leaf N~f~ (%)",
                        paste0("Leaf ", "$\\delta$","^13^C"," (\u2030)")) %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

# Acknowledgements
We thank Burhan Amiji for his outstanding technical assistance. We would like to thank the many Hawkesbury Institute for the Environment staff and students who helped during the experimental harvest.
